### 2011-02-11 doc/MVSeg.tex
Added a few things to the DT section

\subsection{Diffusion tensor primer} A diffusion tensor (DT) is a
symmetric positive definite matrix.  Diffusion tensor imaging derives
a dense field of DT estimates at each position in a three-dimensional
imaging volume by measuring the amount of magnetically-induced water
diffusion in each of $N$ directions.  Directions, here, refers to
positions on the sphere (e.g. longitude and latitude but only in the
northern hemisphere).  Each tensor may be described uniquely as
$$D=\sum_i \lambda_i e_i^T e_i $$ where $e_i$ is a unit 3-vector and
$\lambda_i$ is a scalar in 0 to 1.  The $\lambda_i$ defines the
relative amount of diffusion along the axis defined by $e_i$.  A
highly anisotropic tensor will have $\lambda_1 >> \lambda_{2,3} $, that
is, the prinicipal eigenvector of the tensor will have much higher
diffusion than the other eigenvectors.   The variance of the
$\lambda_i$ is known as the fractional anisotropy (FA) and their mean is
known as the mean diffusion, apparent diffusion coefficient (ADC) or
the trace (actually the sum). Below are some of other measurements derived from DTI:
\begin{description}
\item[RBG:] The principal direction of the tensor can be visualized in
  RGB space.  Some segmentation methods rely on this feature 
  modulated by the FA.  That is, the RBG image voxel may be defined as
  $\vec{rgb}=[R=255~\text{FA}~e_1^1, G=255~\text{FA}~e_1^2, B=255~\text{FA}~e_1^3$].
\item[FA:] Fractional anisotropy derived from the tensor.  This is
  equal to the variance of the eigenvalues of the tensor matrix.  It
  is the most common measure used in DTI studies, along with
  tractography. 
\item[Tensor shape:]  This is determined by the tensor eigenvalues.  As
  neuroscientists interested in tracts, we may be concerned with how
  collections of eigenvalues are distributed along tracts and how
  these distributions differentiate tracts or tissues.
\item[Tensor orientation:] The principal
  eigenvector can point to any position on the upper hemisphere.  As
  communicated by the power of the RGB display of tensor information,
  this is a powerful feature. 
\item[Tractography:] White matter tracts provide a physical connection
  between and within cortical regions.  At the coarse scale of DTI, we mostly
  look at connections between cortical regions.  A simple way to
  define a tract from DTI is to trace a curve that follows the principal eigenvector until
  some stopping criterion based on curvature or FA is reached. 
\end{description}
So, in this work, we'd like to use these different features to help us
automatically differentiate tissues and tracts in not only population
average images but also individual images.  Being able to compute
these types of measures will let us do specific, novel studies of
neural architecture across species, gender, age and disease.

>>>>

\subsection{Diffusion tensor primer} A diffusion tensor (DT) is a
symmetric positive definite matrix.  Diffusion tensor imaging derives
a dense field of DT estimates at each position in a three-dimensional
imaging volume by measuring the amount of water
diffusion along each of $N$ directions. Directions, here, refers to
undirected axes on the unit sphere - a measurement with gradient direction $\vec{x}$ is equivalent to a measurement with
gradient direction $-\vec{x}$. Because of this symmetry, we can describe any measurement
using only positions on the unit hemisphere. (e.g., longitude and latitude can describe any axis through the Earth's center, using only 
latitudes in the northern hemisphere).

Each tensor may be described uniquely as
$$D= \frac1{\mathrm{Tr}(D)}\sum_i \lambda_i e_i^T e_i $$ where $e_i$ is a unit 3-vector and
$\lambda_i$ is a scalar in 0 to 1.  The $\lambda_i$ defines the
relative amount of diffusion along the axis defined by $e_i$.  A
highly anisotropic tensor will have $\lambda_1 >> \lambda_{2,3} $, that
is, the principal eigenvector of the tensor will have much higher
diffusion than the other eigenvectors.   The variance of the
$\lambda_i$ is known as the fractional anisotropy (FA) and their mean is
known as the mean diffusion or apparent diffusion coefficient (ADC). Sometimes you will see the
tensor trace (sum of the eigenvalues) reported instead of the ADC. Below are some of other measurements derived from DTI:
\begin{description}
\item[RGB:] The principal direction of the tensor can be visualized in
  RGB space.  Some segmentation methods rely on this feature 
  modulated by the FA.  That is, the RGB image voxel may be defined as
  $\vec{rgb}=[R=255~\text{FA}~e_1^1, G=255~\text{FA}~e_1^2, B=255~\text{FA}~e_1^3$].
\item[FA:] Fractional anisotropy derived from the tensor.  This is
  equal to the normalized variance of the eigenvalues of the tensor matrix.  It
  is the most common measure used in DTI studies, along with
  tractography. 
\item[Tensor shape:]  This is determined by the tensor eigenvalues.  As
  neuroscientists interested in tracts, we may be concerned with how
  collections of eigenvalues are distributed along tracts and how
  these distributions differentiate tracts or tissues. FA is one shape measure, you can define others to describe
  things like linearity ( $\lambda_1 >> \lambda_{2,3} $) vs planarity ( $\lambda_1 \simeq \lambda_2 >> \lambda_3 $)
  of the ellipsoid tensor representation.
\item[Tensor orientation:] The principal
  eigenvector can point to any position on the upper hemisphere.  As
  communicated by the power of the RGB display of tensor information,
  this is a powerful feature. 
\item[Tractography:] White matter tracts provide a physical connection
  between and within cortical regions.  At the coarse scale of DTI, we mostly
  look at connections between cortical regions.  A simple way to
  define a tract from DTI is to trace a curve that follows the principal eigenvector until
  some stopping criterion based on curvature or FA is reached. 
\end{description}
So, in this work, we'd like to use these different features to help us
automatically differentiate tissues and tracts in not only population
average images but also individual images.  Being able to compute
these types of measures will let us do specific, novel studies of
neural architecture across species, gender, age and disease.
